### 2020-03-28 LaTeX_Code/Discussion/Discussion_14_06_19.tex
I cleaned a bit the directories

\documentclass[a4paper, 12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[activeacute, english]{babel}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{tikz}
\usepackage{float}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{subcaption}

\newcommand\dist{\stackrel{\mathclap{\footnotesize\mbox{d.}}}{=}}
\usepackage[spanish,onelanguage,ruled,lined, linesnumbered]{algorithm2e}
\usepackage[spanish,onelanguage]{algorithm2e}
\newtheorem{theorem}{Teorema}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposición}
%\newenvironment{proof}[1][Demostración]{\begin{trivlist}
%		\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{definition}[1][Definición]{\begin{trivlist}
		\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Ejemplo]{\begin{trivlist}
		\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Nota]{\begin{trivlist}
		\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newcommand{\negr}[1]{\textit{\textbf{#1}}}

\title{Power Forecasting and Control\\ discussion notes}

%\author{Waleed Alhaddad }

\date{}
\begin{document}
\maketitle

\section{Notes on current work flow}
\begin{itemize}
\item Fit and simulate  the model for shorter periods of 6hr, 24hr and compare parameters and behavior.
\item Fit and simulate model without the term $p_t(1-p_t)$  and compare parameters and behavior. Check if it is necessary.
\item Find a way to  prove that the optimization problem is indeed convex in the parameters of the SDE if possible.
\item Reasoning why the sensitivity in the parameter $\alpha$ is much higher than the parameter $\theta$, ellipse suggests that a linear combinations of the two parameters $\alpha, \theta_0$ seem to be equivalent. Visually, we can also  see by trying out parameters, that certain combination of $\alpha$ and $\theta_0$ seem to be indeed very similar to other combinations.
\end{itemize}

\section{Notes on future work}
\begin{itemize}
\item Consider regime switching to detect human energy management decisions.
\end{itemize}

%\newpage
%\section{Things to do}
%\begin{itemize}
%\item 
%\end{itemize}
%
%
%
%\section{Things to do}
%\begin{itemize}
%\item 
%\end{itemize}

%\section{Motivation}

%\section{Introduction}

%\section{Model}

%\section{Conclusion}

%\subsection{Model}
%\begin{thebibliography}{99}
%	\bibitem{apo} Apostol, T. (1974). \emph{Mathematical Analysis} (2nd ed.),  Addison-Wesley. pp.218-230
%	\bibitem{E} Cabaña, E. (2012). \emph{Introducción a los
%		procesos estocásticos}, Notas
%	para el Curso de la Licenciatura en Estadística, Universidad de la República. pp.98-115
%\end{thebibliography}

\end{document}