### 2020-07-06 LaTeX_Code/Annals of Applied Statistics Guidelines/Annals of Applied Statistics Guidelines.tex
We ordered the files. We need to addapt the paper to the jurnals template

\documentclass[12pt]{article}
\usepackage[table]{xcolor}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb}
\usepackage[english]{babel}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{blkarray}
\usepackage{float}
\usepackage{bm}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{commath}

\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{proposition}
\newtheorem{exmp}{Example}[section]\newtheorem{definition}{Definition}[section]
\newtheorem{remark}{Remark}
\newtheorem{ex}{Exercise}
\theoremstyle{definition}
\theoremstyle{remark}
\bibliographystyle{elsarticle-num}

\DeclareMathOperator{\sinc}{sinc}
\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\matindex}[1]{\mbox{\scriptsize#1}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\K}{\mathbb{K}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\prob}{\mathbb{P}}

\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}

%\setcounter{secnumdepth}{0} % This line removes the numbers in the titles.

\begin{document}
\title{Annals of Applied Statistics Guidelines}
\author{Renzo Miguel Caballero Rosas\\
\url{Renzo.CaballeroRosas@kaust.edu.sa}\\
\url{CaballeroRenzo@hotmail.com}\\
\url{CaballeroRen@gmail.com}} 
\maketitle

\tableofcontents

\section{Manuscript Submission}

The abstract and introduction need to say \textbf{What scientific/policy question(s)} the paper addresses, \textbf{Why the topic is of interest}, and \textbf{How the data analyses will address these questions}, in as non-technical a manner as possible.\\
\quad\\
References should be in the name/year format (e.g., Smith, Jones, and Roberts (1992)), using “et al.” for more than three authors; subsequent references can be abbreviated (e.g., Smith et al. (1992)).\\
\quad\\
We encourage use of the IMS LaTex template when preparing your manuscript. The default setting is for numbered citations and references. To get the name-year format, change the line
\begin{itemize}

\item RequirePackage[numbers]\{natbib\} to
\item RequirePackage\{natbib\},

\end{itemize}
i.e., delete [numbers].\\
\quad\\
Authors may use British or American spelling, as they prefer.

\end{document}