### 2020-08-05 Reports/20200330 - Histograms/20200330 - Histograms.tex

\PassOptionsToPackage{table}{xcolor}
\documentclass[aspectratio=169]{beamer}\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage{color}
\usepackage{amsmath,mathtools}
\usepackage{booktabs}
\usepackage{mathptmx}
\usepackage[11pt]{moresize}
\usepackage{hyperref}
\usepackage{commath}
\usepackage{bm}
\usepackage{subfigure}
\usepackage{siunitx}

\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=5mm,text margin right=5mm}
\setbeamertemplate{caption}[numbered]
\addtobeamertemplate{navigation symbols}{}{
\usebeamerfont{footline}
\usebeamercolor[fg]{footline}
\hspace{1em}
\insertframenumber/\inserttotalframenumber}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\K}{\mathbb{K}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\I}{\mathbb{I}}
\DeclareMathOperator{\sign}{sign}

\title{Data Histograms}
\subtitle{Renzo Miguel Caballero Rosas}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Briefly description}

In the following slides, we will show histograms for the error ($V_{t_i}=X_{t_i}-p_{i_i}$), and the Lamperti transform ($Z_{t_i}=\psi(V_{t_i})$). For each one, we will show the transitions and the real measurements.\\
\quad\\
Also, for each histogram, we will find its empirical mean and variance, and approximate it with a Gaussian distribution.

\end{frame}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Error transitions and Lamperti transitions}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.75\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx.eps}
\end{figure}

\end{frame}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Error transitions and Lamperti transitions}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.48\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx_Err.eps}\quad
\includegraphics[width=0.48\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx_Lam.eps}
\end{figure}

\end{frame}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Error measurements and Lamperti measurements}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.75\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx_Measurements.eps}
\end{figure}

\end{frame}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Error measurements and Lamperti measurements}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.48\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx_Measurements_Error.eps}\quad
\includegraphics[width=0.48\textwidth]{../../../Python/Represas_Data_2/Wind_Data/someResults/final/Gauss_Approx_Measurements_Lamperti.eps}
\end{figure}

\end{frame}

\end{document}